---
title: "t, f , Chi-square tests and ANOVA"
author: "INSOFE"
date: "22 December 2018"
output:
  html_document: 
      toc: yes
      toc_float : yes
      toc_depth : 3
  
---

**NOTE** Before starting this assignment please remember to clear your environment, you can do that by running the following code chunk

```{r}

rm(list = ls(all=TRUE))

```

## Agenda 

* Type I and Type II error

* t-test

* Chi-square test

* f-test

* ANOVA

## Z-test
### Problem 1

* Suppose a manufacturer claims that the mean lifetime of a light bulb is at least 10,000 hours. In a sample of 30 light bulbs, it was found that they only last 9,900 hours on average. Assuming the population standard deviation to be 120 hours, at 0.05 significance level, can we reject the claim by the manufacturer?

```{r}

# Please refer "20181216_Batch56_CSE7315c_Lab05_Ztest_ttest_Solutions.pdf" for solution

```

## t-test
### Problem 2
* Suppose a car manufacturer claims a model gets at least 25 mpg. A consumer group asks 10 owners of this model to calculate their mpg and the mean value was 22 with a standard deviation of 1.5. Is the manufacturer's claim supported at 95%confidence level.

```{r}

rm(list=ls(all=TRUE))

# H0: mu >= 25, H1: mu < 25. 

n = 10

xBar = 22 

mu = 25

s = 1.5

alpha = 0.05

# Since sample size is small, we will use Student's t distribution here.  This is lower tailed test.

se = s/sqrt(n)

test_Statistic = (xBar - mu)/ se

degrees_Of_Freedom = n-1 

tValue= qt(alpha, degrees_Of_Freedom)
test_Statistic
tValue

# We reject the NUll Hypothesis since the observation falls in rejection/critical region.

```

### Problem 3
* An outbreak of Salmonella-related illness was attributed to ice cream produced at a certain factory. Scientists measured the level of Salmonella in 9 randomly sampled batches of ice cream. The levels (in MPN/g) were: 0.593 0.142 0.329 0.691 0.231 0.793 0.519 0.392 0.418. Is there evidence that the mean level of Salmonella in the ice cream is greater than 0.3 MPN/g.

```{r}

rm(list=ls(all=TRUE))
# One-sample t-tests
# H0 <= 0.3 vs H1 > 0.3 
Salmonella_level = c(0.593, 0.142, 0.329, 0.691, 0.231, 0.793, 0.519, 0.392, 0.418)

######### t-test ##########################
# General form 
#t.test(x, y = NULL, alternative = c("two.sided", "less", "greater"), mu = 0, paired = FALSE, var.equal = FALSE, conf.level = 0.95)

t.test(Salmonella_level, alternative="greater", mu=0.3)
tValue= qt(0.95, 8)
tValue

# Since test statistic value t = 2.2051 > tValue, we reject the null hypothesis. 
```

### Problem 4
* A study was performed to test whether cars get better mileage on premium gas than on regular gas. Each of 10 cars was first filled with either regular or premium gas, decided by a coin toss, and the mileage for that tank was recorded. The mileage was recorded again for the same cars using the other kind of gasoline. Test to determine whether cars get significantly better mileage with premium gas	 	
Reg : 16, 20, 21, 22, 23, 22, 27, 25, 27, 28 
Prem : 19, 22, 24, 24, 25, 25, 26, 26, 28, 32

```{r}
rm(list=ls(all=TRUE))
# H0: Mileage on premium gas <= Mileage on regular gas
# H1: Mileage on premium gas > Mileage on regular gas

# A study was performed to test whether cars
# Below is the relevant R-code: 
prem = c(19, 22, 24, 24, 25, 25, 26, 26, 28, 32) 
reg = c(16, 20, 21, 22, 23, 22, 27, 25, 27, 28) 

t.test(prem,reg,alternative="greater", paired=TRUE)

tValue= qt(0.95, 9)
tValue

# Since test statistic value t > tValue, we reject the null hypothesis. 

```

## Chi-square test
### Problem 5
* Do people have a preference for movie type?
There are four categories of movies: comedy, horror, drama and science fiction. Let us say, we assumed a simple uniform distribution, where each category is liked by 25% of the people. Now we need to test how good our guess is:

* Observed Frequency
- Comedy - 35
- Horror - 30
- Drama - 20
- Science Fiction - 15
- Total (n) = 100

```{r}

# H0: the observed distribution fits the expected 
# H1: the observed distribution does not fit the expected 

rm(list = ls(all=T))
observed <- matrix(c(35, 30,20,15), ncol = 1)
observed
expected <- rep(sum(observed)/4,4)
expected
test_stat<-  sum((observed - expected)^2 / expected)
print(test_stat)

critvalue<- qchisq(0.05, 3, lower.tail = F)
print(critvalue)

test_stat1=chisq.test(observed)
test_stat1$expected

test_stat1

# Since test statistic value test_stat > critvalue, hence, we reject the Null Hypothesis.

```

### Problem 6
* A survey is conducted by a gaming company that makes three video games. It wants to know if the preference of game depends on the gender of the player. Total number of participants is 1000. 

a. State the null hypothesis and alternate hypothesis.
b. Calculate the degrees of freedom.
c. Does men's preference is different from women's preference? Check with 0.05 level of significance.

```{r}

rm(list = ls(all=T))
# Null hypothesis H0: the preference of game is independent of gender 
# Alternate hypothesis H1: the preference of game is dependent on gender

# If a contingency table of observed values is given, chisq. test does an independence test
observed <- matrix(c(200,150,50,250,300,50), byrow = TRUE, ncol = 3)

test_stat <- chisq.test(observed)
test_stat

test_stat$expected

crit <- qchisq(p = 0.05, df = (nrow(observed)-1)*(ncol(observed)-1) ,lower.tail = FALSE)
crit

# critical region lies at p(x2 > 5.991)

# test_stat$statistic > crit

# Hence we reject NULL Hypothesis

```

## f-test
### Problem 7

* Laptop computer maker uses battery packs supplied by two companies, A and B. While both brands have the same average battery life between charges (LBC), the computer maker seems to receive more complaints about shorter LBC than expected for battery packs supplied by company B. The computer maker suspects that this could be caused by higher variance in LBC for Brand B. To check that, ten new battery packs from each brand are selected, installed on the same models of laptops, and the laptops are allowed to run until the battery packs are completely discharged. The following are the observed LBCs in hours.

Brand A = 3.2, 3.4, 2.8, 3, 3, 3, 2.8, 2.9, 3, 3
Brand B = 3, 3.5, 2.9, 3.1, 2.3, 2, 3, 2.9, 3, 4.1
Test at the 10% level of significance, whether the variance of both the brands are similar.

```{r}

# H0:	Var(A) = Var(B)
# H1:	Var(A) != Var(B)

obsA = c(3.2,3.4,2.8,3,3,3,2.8,2.9,3,3)

obsB = c(3,3.5,2.9,3.1,2.3,2,3,2.9,3,4.1)

sdA = sd(obsA)
sdB = sd(obsB)

Fstat = ((sdA)^2)/((sdB)^2)
Fstat

qf(p = 0.05,df1 = 9,df2 = 9)
qf(p = 0.95,df1 = 9,df2 = 9)

# Variances of both the brands are not similar.

```

## ANOVA
### Problem 8
* A car crash research team wants to examine the safety of compact cars, intermediate and full size cars. Given below are the hypothetical values of the mean pressure applied to the drivers head during the crash test for each of the car types. Check whether means are equal for each type of these cars at 5% significance level.

```{r}

# H0: The mean pressure applied to the driver heads for each of the groups are the same. 
# H1: Atleast one of the mean pressure applied to the driver heads for the groups are different

x1 <- c(643, 655, 702)
x2 <- c(469, 427, 525)
x3 <- c(484, 456, 402)

x <- c(x1,x2,x3)
x

sst <- sum((x- mean(x))^2)

sst

m <- 3

n <- length(x1)
n

ssw <- sum((x1- mean(x1))^2) + sum((x2- mean(x2))^2) + sum((x3- mean(x3))^2)

ssw

ssb <- sst - ssw

ssb

df_ssw <- m * (n-1)
df_ssb <- m-1

f_stat <- (ssb / df_ssb) / (ssw / df_ssw)

f_stat

F_crit <- qf(0.05, df_ssb, df_ssw, lower.tail = F)
F_crit
F_crit1 <- qf(0.95, df_ssb, df_ssw)
F_crit1

# F-statistic > F-critical; Hence the statistic lies in the critical region.
# We fail to accept H0; Reject H0

```



