---
title: "Analysis & Workflow Arrangement"
output: html_document
---

# Field Work {.tabset}

## Objective

```{r}
print("Client provides the requirement as well as the Objective")
```

## Domain Understanding 

```{r}
print("Do the Domain Understanding before even starting the Data Analysis")
```

## Architecting

```{r}
print("Architecting the Solution Work Flow i.e What needs to be done for achieving the goal")
```

## Evaluation Metric

```{r}
print("What is my my Evaluation Metric, with which i can evaluate my Modelling & Predictions")
```

# Exploratory Data Analysis (EDA) {.tabset}

## Str & Summ

```{r}
print("Understanding the Structure & Summary of the given Data")
```

## NA's

```{r}
print("Identifying the Missing values in the Dataset")
```

## Other Data Pre-Processing Techniques

```{r}
print("Identifying the Missing values in the Dataset")
```

## Visualization

```{r}
# print("Visualizing the Data to extract insights from it")
```

### Various Sections in the Visualization {.tabset}

#### Uni-variate Plots

```{r}
print("Generate your Uni-variate Plots here")
```

#### Bi-variate Plots

```{r}
print("Generate your Bi-variate Plots here")
```

#### Multi-variate Plots

```{r}
print("Generate your Multi-variate Plots here")
```

# Data_Modelling

## Basic Models

```{r}
print("Build your basic model to set a threshold - To build models that should be better than this")
```

## Feature Importance

```{r}
print("Identifying the Feature importance using various Feature Techniques - Step-AIC/VIF/RF")
```

## Advance Models {.tabset}

```{r}
print("Generate your Multi-variate Plots here")
```

### Decision Trees

```{r}
print("Build your Decision Tree Algo")
```

### Naive Bayes 

```{r}
print("Build your Naive Bayes Algo")
```

### K-Nearest Neighbors

```{r}
print("Build your KNN Algo")
```

### Support Vector Machines

```{r}
print("Build your SVM Algo")
```

### Ensemble Learning {.tabset}

```{r}
# print("Build your Ensemble Models")
```

#### Bagging

```{r}
print("Bagging Model")
```

#### Boosting

```{r}
print("Boosting Model")
```

#### Stacking

```{r}
print("Stacking Model")
```

# Evaluation of various Models {.tabset}

## Decision Trees Confusion Matrix

```{r}
print("Generate your Decision Tree Algo's Confusion Matrix")
```

## Naive Bayes Confusion Matrix

```{r}
print("Generate your Naive Bayes Algo's Confusion Matrix")
```

## K-Nearest Neighbors Confusion Matrix

```{r}
print("Generate your KNN Algo's Confusion Matrix")
```

## Support Vector Machines Confusion Matrix

```{r}
print("Generate your SVM Algo's Confusion Matrix")
```

## Ensemble Learning {.tabset}

```{r}
# print("Generate your Ensemble Models CM's")
```

### Bagging Confusion Matrix

```{r}
print("Generate your Bagging Model's Confusion Matrix")
```

### Boosting Confusion Matrix

```{r}
print("Generate your Boosting Model's Confusion Matrix")
```

### Stacking Confusion Matrix

```{r}
print("Generate your Stacking Model's Confusion Matrix")
```

# End-Point

```{r}
# if(results == satisfactory){
#   print("Converge & Submit the results")
# }
# else{
#   print("Re-Iterate over the same work flow by tweaking some more aspects of the work flow")
# }
```
  

# ----------------------- Adios ---------------------------

