---
title: "Tidyverse LabActivity1"
author: "Insofe Lab"
date: "17th Feb 2019"
output:
  html_document: default
  pdf_document: default
---

```{r}
# Removing the global environment variables
rm(list = ls(all = TRUE))
```

```{r}
library(tidyverse)
```

```{r}
# Setting the working directory
# Set your Own Working directory Here if the Dataset & the ".Rmd" file are in different paths
# setwd("")
```
# tidyr package
## separate() method: It splits single column to multiple columns
```{r}
"inc5000_2017.csv" %>% 
  read.csv() %>%
    separate(founded, c("date", "month", "year_founded"), sep="/", remove = TRUE) -> inc5000_2017_data 
head(inc5000_2017_data)  
```
## unite() method: It combines multiple columns to a single column
```{r}
inc5000_2017_data %>%
  unite(founded, c(date,month,year_founded), sep= "/", remove = TRUE) -> inc5000_2017_data2
head(inc5000_2017_data2)
```

## spread() method: Used for converting data from long format to wide format
```{r}
"marklist_long.csv" %>% 
  read.csv() %>%
    spread(key= Subject, value = Marks) -> wide_data

print(wide_data)
```

## gather() method: Used for converting data from wide format to long format
```{r}
"marklist_wide.csv" %>%  
  read.csv() %>% 
    gather(Subject, Marks, c(English,Maths,Science)) -> long_data
print(long_data)
```

# dplyr package
## select() method: To pick specific columns
```{r}
"marklist_long.csv" %>%  
read_csv() %>% 
select(Name,Subject)  
```
## filter() method: To filter specific rows based on some condition
```{r}
"marklist_long.csv" %>%  
read_csv() %>% 
filter(Name == "Ram")
```
## arrange() method: To sort the rows
```{r}
"marklist_long.csv" %>%  
read_csv() %>% 
arrange(desc(Marks))
```
## mutate() method: To create new variables that are functions of existing variables
```{r}
"marklist_long.csv" %>%  
read_csv() %>% 
mutate(grade= factor((Marks >= 80), labels = c("Passed", "Excellence")))
```
## summarize() and group_by() methods: To group the columns and summarize the results
```{r}
"marklist_long.csv" %>%  
read_csv() %>% 
group_by(Name) %>%
summarize(TotalMarks = sum(Marks))
```




